LLM-Driven CIDOC-CRM Knowledge Graph Extraction from Museum Archives

### 2025-05-04 paper/paper.tex
paper: add impact column to graphs
paper: revise IV. Results

\begin{table}[h]
    \centering
    \caption{KBQA score (overall)}
    \label{tab:qa_percent}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{60.50}                   \\
        \hline
        Qwen 2.5 3B & 73.50                       & 68.50           \\
        \hline
        Llama 3 3B  & 79.50                       & 78.00           \\
        \hline
        Qwen 2.5 7B & 95.50                       & 87.50           \\
        \hline
        Llama 3 8B  & 92.00                       & 90.50           \\
        \hline
        DeepSeek V3 & \textbf{98.00}              & \textbf{97.50}  \\
        \hline
        DeepSeek R1 & 97.50                       & 95.00           \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (single item)}
    \label{tab:qa_percent_single}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{62.16}                   \\
        \hline
        Qwen 2.5 3B & 72.97                       & 67.57           \\
        \hline
        Llama 3 3B  & 80.54                       & 78.92           \\
        \hline
        Qwen 2.5 7B & 95.68                       & 89.19           \\
        \hline
        Llama 3 8B  & 94.05                       & 91.89           \\
        \hline
        DeepSeek V3 & \textbf{100}                & \textbf{98.38}  \\
        \hline
        DeepSeek R1 & 98.92                       & 96.22           \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (multi item)}
    \label{tab:qa_percent_multi}
    \begin{tabular}{|l|c|c|}
        \hline
        Graph       & Ontology-Free            & Ontology-Guided \\
        \hline
        Blank       & \multicolumn{2}{|c|}{40}                   \\
        \hline
        Qwen 2.5 3B & 80                       & 80              \\
        \hline
        Llama 3 3B  & 66.67                    & 66.67           \\
        \hline
        Qwen 2.5 7B & \textbf{93.33}           & 66.67           \\
        \hline
        Llama 3 8B  & 66.67                    & 73.33           \\
        \hline
        DeepSeek V3 & 73.33                    & \textbf{86.67}  \\
        \hline
        DeepSeek R1 & 80                       & 80              \\
        \hline
    \end{tabular}
\end{table}

The results of the Knowledge Base Question Answering (KBQA) task, designed to evaluate the practical utility of the generated knowledge graphs, are presented in Tables \ref{tab:qa_percent}, \ref{tab:qa_percent_single}, and \ref{tab:qa_percent_multi}. These tables report the overall accuracy, the accuracy on single-item questions, and the accuracy on multi-item questions, respectively, for both ontology-free and ontology-guided knowledge graphs generated by each tested LLM, in addition to the no-context "blank" test.

A notable observation is the relatively high performance of the "Blank" baseline across all question types, particularly on the single-item questions (62.16\% accuracy). This suggests that the LLMs, even without access to the generated knowledge graphs, possess a significant amount of information relevant to the Museum Nasional collection, potentially acquired during their pretraining. It is also possible that a subset of the questions could be answered through general reasoning capabilities without specific knowledge of the museum's data. The lower performance of the "Blank" baseline on multi-item questions (40\% accuracy) indicates that these more complex queries necessitate the integration of information, which the models struggle with in the absence of context.

The general trend on KBQA results is that larger models tend to yield better KBQA results. This aligns with the findings from the ontology conformance evaluation, suggesting that increased model capacity contributes to improved information extraction and reasoning capabilities. However, compared to ontology conformance, the differences are not so extreme. This suggests that while smaller models struggle to reason around ontology constraints, they still have reasonable capability to extract information.

Comparing the ontology-free and guided approaches, while ontology-free graphs achieve marginally higher scores, the overall difference remains small. This close performance between ontology-free and ontology-guided graphs has important implications. It suggests that while enforcing ontological structure does not impede the extraction of answerable information, it also doesn't provide a substantial boost in question-answering accuracy, at least in this specific evaluation setting. This could be due to several factors, including the nature of the questions asked, the effectiveness of the ontology integration in the prompting process, or the inherent ability of the LLMs to extract relevant information even without explicit ontological guidance.

>>>>

\begin{table}[h]
    \centering
    \caption{KBQA score (overall)}
    \label{tab:qa_percent}
    \begin{tabular}{|l|c|c|r|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided & Impact \\
        \hline
        Blank       & \multicolumn{2}{|c|}{60.50} &                          \\
        \hline
        Qwen 2.5 3B & 73.50                       & 68.50           & -5.00  \\
        \hline
        Llama 3 3B  & 79.50                       & 78.00           & -1.50  \\
        \hline
        Qwen 2.5 7B & 95.50                       & 87.50           & -8.00  \\
        \hline
        Llama 3 8B  & 92.00                       & 90.50           & -1.50  \\
        \hline
        DeepSeek V3 & \textbf{98.00}              & \textbf{97.50}  & -0.50  \\
        \hline
        DeepSeek R1 & 97.50                       & 95.00           & -2.50  \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (single item)}
    \label{tab:qa_percent_single}
    \begin{tabular}{|l|c|c|r|}
        \hline
        Graph       & Ontology-Free               & Ontology-Guided & Impact \\
        \hline
        Blank       & \multicolumn{2}{|c|}{62.16} &                          \\
        \hline
        Qwen 2.5 3B & 72.97                       & 67.57           & -5.40  \\
        \hline
        Llama 3 3B  & 80.54                       & 78.92           & -1.62  \\
        \hline
        Qwen 2.5 7B & 95.68                       & 89.19           & -6.49  \\
        \hline
        Llama 3 8B  & 94.05                       & 91.89           & -2.16  \\
        \hline
        DeepSeek V3 & \textbf{100.00}             & \textbf{98.38}  & -1.62  \\
        \hline
        DeepSeek R1 & 98.92                       & 96.22           & -2.70  \\
        \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \centering
    \caption{KBQA score (multi item)}
    \label{tab:qa_percent_multi}
    \begin{tabular}{|l|c|c|r|}
        \hline
        Graph       & Ontology-Free            & Ontology-Guided & Impact \\
        \hline
        Blank       & \multicolumn{2}{|c|}{40} &                          \\
        \hline
        Qwen 2.5 3B & 80.00                    & 80.00           & 0.00   \\
        \hline
        Llama 3 3B  & 66.67                    & 66.67           & 0.00   \\
        \hline
        Qwen 2.5 7B & \textbf{93.33}           & 66.67           & -26.66 \\
        \hline
        Llama 3 8B  & 66.67                    & 73.33           & 6.66   \\
        \hline
        DeepSeek V3 & 73.33                    & \textbf{86.67}  & 13.34  \\
        \hline
        DeepSeek R1 & 80.00                    & 80.00           & 0.00   \\
        \hline
    \end{tabular}
\end{table}

The results of the Knowledge Base Question Answering (KBQA) task, designed to evaluate the practical utility of the generated knowledge graphs, are presented in Tables \ref{tab:qa_percent}, \ref{tab:qa_percent_single}, and \ref{tab:qa_percent_multi}. These tables report the overall accuracy (n=200), the accuracy on single-item questions (n=185), and the accuracy on multi-item questions (n=15), respectively, for both ontology-free and ontology-guided knowledge graphs generated by each tested LLM, in addition to the no-context "blank" test. Providing any graph as context significantly improved QA performance over the blank baseline, confirming the value of extracted knowledge. The blank baseline's reasonable performance, especially on single-item questions, suggests the evaluator model (DeepSeek V3) possesses some inherent knowledge related to the domain.

Consistent with the ontology conformance results, larger models generally produced graphs that lead to better KBQA scores, although the performance gap between models was less pronounced here than in the conformance task. This indicates that even models struggling with strict ontology adherence can still extract useful factual information.

Comparing the generation approaches, ontology-free graphs surprisingly led to slightly better or equal overall and single-item KBQA performance across all models. This suggests that for retrieving specific facts, the added structure of the ontology-guided graphs did not translate into improved accuracy in this setup, and may have even slightly hindered it in some cases.

The results for multi-item questions in particular were more varied. Some models benefited from ontology guidance, while others performed worse or showed no difference. However, conclusions drawn from this subset should be cautious due to the small number of multi-item questions (n=15). A larger, more diverse set of complex questions would be needed to reliably assess the impact of ontology guidance on reasoning across the graph.

Overall, the KBQA results show that LLM-generated graphs enhance question answering, and larger models tend to perform better. However, no benefit from ontology guidance during generation for KBQA performance was consistently observed, and requires further investigation with a larger and more refined set of questions.
